Texture mapping 3D planar models of indoor environments with noisy camera poses

### 2012-12-25 spie/paper.tex
going oing going

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{abstract}
  Automated 3D modeling of building interiors is used in applications
  such as virtual reality and environment mapping. Incorporating
  images and texture mapping these models allows for photorealistic
  visualizations of the data collected by such modeling systems. Pose
  estimates obtained by these systems often suffer from inaccuracies,
  resulting in visible discontinuities when successive images are
  projected adjacently onto a surface for texturing. Existing methods
  to stitch images together are often computationally expensive and
  work independently of pose estimates and geometry information. We
  propose a method to refine camera poses using both existing
  estimates and geometry information, followed by two different
  methods to composite images together, based on the uniformity of
  available images. The effectiveness of our methods are demonstrated
  on a number of disparate indoor environments.
\end{abstract}

@@

The first step in automated 3D modeling is the physical scanning of
the environment's geometry. An indoor modeling system must recover its
own poses within an environment while simultaneously reconstructing
the 3D structure of the environment itself. CITATION. This is known as
the simultaneous localization and mapping (SLAM) problem, and is
generally solved by taking readings from laser range scanners,
cameras, and inertial measurement units (IMUs) at multiple locations
within the environment.

In this paper, we work with data obtained from a backpack-mounted
system, carried by an ambulatory human. Such a system provides
advantages over more common wheel-mounted systems in terms of agility
and portability, but results in much higher localization error and
uncertainty overall. CITATION. As a result, common methods for texture
mapping generally produce poor results, leading to the development of
the approaches contained in this paper. Besides localization
information and images, environment geometry is required as well for
texture mapping. In this paper we work mainly with models obtained
through plane-fitting surfaces to point clouds generated by the
backpack system. CITATION. We have also developed an appropriate
method to work with near-planar models, as described in APPENDIX
SOMETHING.

Though our texture mapping procedure was designed with the
aforementioned system in mind, it is easily generalizable to any other
system, provided the required inputs of surface geometry, images with
rough extrinsic pose estimates, and either provided or estimated
intrinsic camera calibration matrices. An added benefit of our method
is its relatively fast running time and modularity of each step,
allowing for quick tuning of parameters and easy accomodation of new
input data given updates in localization or model-generation
algorithms.

The remainder of the paper is organized as follows. Section X explains
how input images are projected onto our geometry and simple approaches
towards texturing. Section X covers more sophisticated existing
approaches to image matching and stitching, and their performance on
our datasets. Section X contains our approach towards image matching,
followed by Section X, which describes compositing approaches. Section
X contains further results and conclusions.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsection{Direct Mapping}
\label{sec:directMapping}

@@

Once we have a list of candidate images for $t$, we define a scoring
function in order to objectively select the best image for texturing
$t$. Since camera pose errors compound over distance, we wish to
minimize the distance between cameras and the surfaces they
texture. Additionally, we desire images that are projected
perpendicularly onto the plane, maximizing the resolution and amount
of useful texture available in their projections, as well as
minimizing any parallax effects due to real-world geometry not
represented by our digital model. In other words, we wish to minimize
the angle between the tile's normal vector and the camera axis for
images selected for texture mapping. These two criteria can be met by
maximizing the function $\frac{1}{d} (-1 \cdot \vec{c}) \cdot \vec{n}$
as shown in Figure \ref{fig:scoringFunction}. Specifically, $d$ is the
distance between the centers of a camera and a tile, and $\vec{n}$ and
$\vec{c}$ are the directions of the plane's normal and the camera axis
respectively.

@@

As Figure \ref{fig:compareAll}(a) demonstrates, this approach leads to
the best texture for each tile independently, but overall results in
many image boundaries with abrupt discontinuities, due to significant
misalignment between images, as a result of camera pose inaccuracies.

\subsection{Mapping with Caching}
\label{sec:mappingWithCaching}
Since discontinuities occur where adjacent tiles select non-aligned
images, it makes sense to take into account image selections made by
neighboring tiles while texture mapping a given tile. By using the
same image across tile boundaries, we can eliminate a discontinuity
altogether. If this is not possible because a tile is not visible in
images used by neighboring tiles, using similar images across tile
boundaries also leads to less noticeable discontinuities.

@@

\section{Our Image Alignment with Cool Name}
\section{Image Compositing}
\section{Results and Conclusions}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%poop below
\section{PARTS OF MANUSCRIPT}

This section describes the normal structure of a manuscript and how
each part should be handled.  The appropriate vertical spacing between
various parts of this document is achieved in LaTeX through the proper
use of defined constructs, such as \verb|\section{}|.  In LaTeX,
paragraphs are separated by blank lines in the source file.

At times it may be desired, for formatting reasons, to break a line
without starting a new paragraph.  This situation may occur, for
example, when formatting the article title, author information, or
section headings.  Line breaks are inserted in LaTeX by entering
\verb|\\| or \verb|\linebreak| in the LaTeX source file at the desired
location.

%%%%% Sometimes it is necessary to precede the double slash by
%%%%% \verb|\protect| to get the desired result, for example, in
%%%%% article titles.

%% -----------------------------------------------------------
\subsection{Title and Author Information}
\label{sec:title}

The article title appears centered at the top of the first page.  The
title font is 16 point, bold.  The rules for capitalizing the title
are the same as for sentences; only the first word, proper nouns, and
acronyms should be capitalized.  Avoid using acronyms in the title.
Keep in mind that people outside your area of expertise might read
your article.  Appendix~\ref{sec:misc} contains more about acronyms.

The list of authors immediately follows the title after a blank
vertical space of about 7 mm.  The font is 12 point, normal with each
line centered.  The authors' affiliations and addresses follow the
author list after another blank space of about 4 mm, also in 12-point,
normal font and centered.  Do not use acronyms in affiliations and
addresses. For multiple affiliations, each affiliation should appear
on a new line, separated from the following address by a semicolon.
Italicized superscripts may be used to identify the correspondence
between the authors and their respective affiliations.  Further author
information, such as e-mail address, complete postal address, and
web-site location, may be provided in a footnote by using
\verb|\authorinfo{}|, as demonstrated above.

When the abbreviated title or author information is too long to fit on
one line, multiple lines may be used; insert line breaks appropriately
to achieve a visually pleasing format.  The proper spacing of one and
one-half lines between the title, author list, and their affiliations
is achieved with the command \verb|\skiplinehalf| defined in {\tt
  spie.cls}.

%% -----------------------------------------------------------
\subsection{Abstract and Keywords}
The title and author information is immediately followed by the
Abstract. The Abstract should concisely summarize the key findings of
the paper.  It should consist of a single paragraph containing no more
than 200 words.  The Abstract does not have a section number.  A list
of up to ten keywords should immediately follow the Abstract after a
blank line.  These keywords will be included in a searchable database
at SPIE.

%% -----------------------------------------------------------
\subsection{Body of Paper}
The body of the paper consists of numbered sections that present the
main findings.  These sections should be organized to best present the
material.  See Sec.~\ref{sec:sections} for formatting instructions.

%% -----------------------------------------------------------
\subsection{Appendices}
Auxiliary material that is best left out of the main body of the
paper, for example, derivations of equations, proofs of theorems, and
details of algorithms, may be included in appendices.  Appendices are
enumerated with uppercase Latin letters in alphabetic order, and
appear just before the Acknowledgments and References.

%% -----------------------------------------------------------
\subsection{Acknowledgments}
In the Acknowledgments section, appearing just before the References,
the authors may credit others for their guidance or help.  Also,
funding sources may be stated.  The Acknowledgments section does not
have a section number.

%% -----------------------------------------------------------
\subsection{References}
The References section lists books, articles, and reports that are
cited in the paper.  It does not have a section number.  The
references are numbered in the order in which they are cited.
Examples of the format to be followed are given at the end of this
document.

The reference list at the end of this document is created using
BibTeX, which looks through the file {\tt report.bib} for the entries
cited in the LaTeX source file.  The format of the reference list is
determined by the bibliography style file {\tt spiebib.bst}, as
specified in the \verb|\bibliographystyle{spiebib}| command.
Alternatively, the references may be directly formatted in the LaTeX
source file.

For books\cite{Lamport94,Alred03,Goossens97} the listing includes the
list of authors, book title (in italics), page or chapter numbers,
publisher, city, and year of publication.  A reference to a journal
article\cite{Metropolis53} includes the author list, title of the
article (in quotes), journal name (in italics, properly abbreviated),
volume number (in bold), inclusive page numbers, and year.  By
convention\cite{Lamport94}, article titles are capitalized as
described in Sec.~\ref{sec:title}.  A reference to a proceedings paper
or a chapter in an edited book\cite{Gull89a} includes the author list,
title of the article (in quotes), conference name (in italics), if
appropriate, editors, volume or series title (in italics), volume
number (in bold), if applicable, inclusive page numbers, publisher,
city, and year.  References to an article in the SPIE Proceedings may
include the conference name, as shown in Ref.~\citenum{Hanson93c}.

Citations to the references are made using superscript numerals, as
demonstrated in the preceding paragraph.  One may also directly refer
to a reference within the text, e.g., ``as shown in
Ref.~\citenum{Metropolis53} ..."

%% -----------------------------------------------------------
\subsection{Footnotes}
Footnotes\footnote{Footnotes are indicated as superscript symbols to
  avoid confusion with citations.} may be used to provide auxiliary
information that doesn't need to appear in the text, e.g., to explain
measurement units.  They should be used sparingly, however.

Only nine footnote symbols are available in LaTeX. If you have more
than nine footnotes, you will need to restart the sequence using the
command \verb|\footnote[1]{Your footnote text goes here.}|. If you
don't, LaTeX will provide the error message {\tt Counter too large.},
followed by the offending footnote command.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{SECTION FORMATTING} \label{sec:sections}

Section headings are centered and formatted completely in uppercase
11-point bold font.  Sections should be numbered sequentially,
starting with the first section after the Abstract.  The heading
starts with the section number, followed by a period.  In LaTeX, a new
section is created with the \verb|\section{}| command, which
automatically numbers the sections.

Paragraphs that immediately follow a section heading are leading
paragraphs and should not be indented, according to standard
publishing style\cite{Lamport94}.  The same goes for leading
paragraphs of subsections and sub-subsections.  Subsequent paragraphs
are standard paragraphs, with 14-pt.\ (5 mm) indentation.  An extra
half-line space should be inserted between paragraphs.  In LaTeX, this
spacing is specified by the parameter \verb|\parskip|, which is set in
{\tt spie.cls}.  Indentation of the first line of a paragraph may be
avoided by starting it with \verb|\noindent|.

%% -----------------------------------------------------------
\subsection{Subsection Attributes}

The subsection heading is left justified and set in 11-point, bold
font.  Capitalization rules are the same as those for book titles.
The first word of a subsection heading is capitalized.  The remaining
words are also capitalized, except for minor words with fewer than
four letters, such as articles (a, an, and the), short prepositions
(of, at, by, for, in, etc.), and short conjunctions (and, or, as, but,
etc.).  Subsection numbers consist of the section number, followed by
a period, and the subsection number within that section.

%% -----------
\subsubsection{Sub-subsection attributes}
The sub-subsection heading is left justified and its font is 10 point,
bold.  Capitalize as for sentences.  The first word of a
sub-subsection heading is capitalized.  The rest of the heading is not
capitalized, except for acronyms and proper names.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{FIGURES AND TABLES}

Figures are numbered in the order of their first citation.  They
should appear in numerical order and on or after the same page as
their first reference in the text.  Alternatively, all figures may be
placed at the end of the manuscript, that is, after the Reference
section.  It is preferable to have figures appear at the top or bottom
of the page.  Figures, along with their captions, should be separated
from the main text by at least 0.2 in.\ or 5 mm.

Figure captions are centered below the figure or graph.  Figure
captions start with the figure number in 9-point bold font, followed
by a period; the text is in 9-point normal font; for example,
``{\footnotesize{Figure 3.}  Original image...}".  See
Fig.~\ref{fig:example} for an example of a figure caption.  When the
caption is too long to fit on one line, it should be justified to the
right and left margins of the body of the text.

Tables are handled identically to figures, except that their captions
appear above the table.
%% Use following command to specify that graphics file is in a
%% directory other than this LaTeX source file.  Note use of / to
%% separate subdirectories, for UNIX and Windows OS.
%% \graphicspath{{H:/HANSON/SPIESTY/}} tabular environment useful for
%% creating an array of images
% -------------
\begin{figure}
  \begin{center}
    \begin{tabular}{c}
      \includegraphics[height=7cm]{naive.jpg}
    \end{tabular}
  \end{center}
  \caption[example]
  % >>>> use \label inside caption to get Fig. number with \ref{}
  { \label{fig:example} Figure captions are used to describe the
    figure and help the reader understand it's significance.  The
    caption should be centered underneath the figure and set in
    9-point font.  It is preferable for figures and tables to be
    placed at the top or bottom of the page. LaTeX tends to adhere to
    this standard.}
\end{figure}
% -------------

For further details concerning the inclusion of grayscale and color
images, consult SPIE's {\it Author Guide for Publication and
  Presentation}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix    %>>>> this command starts appendixes
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{MISCELLANEOUS FORMATTING DETAILS} \label{sec:misc}

It is often useful to refer back (or forward) to other sections in the article.  Such references are made by section number.  When a section reference starts a sentence, Section is spelled out; otherwise use its abbreviation, for example, ``In Sec.~2 we showed..." or ``Section~2.1 contained a description...".  References to figures, tables, and theorems are handled the same way.

At the first occurrence of an acronym, spell it out, followed by the acronym in parentheses, e.g., noise power spectrum (NPS).

%%-----------------------------------------------
\subsection{Formatting Equations} 
Equations may appear in line with the text, if they are simple, short, and not of major importance; e.g., $\beta = b/r$.  Important equations appear on their own line.  Such equations are centered.  For example, ``The expression for the minus-log-posterior is
	\begin{equation}
	\label{eq:alpha}
\phi = |{\rm\bf y} - {\rm\bf A}{\rm\bf x}|^2 + \alpha \log p({\rm\bf x}) \, ,
	\end{equation}
where $\alpha$ determines the strength of ..."  Principal equations are numbered, with the equation number placed within parentheses and right justified.

Equations are considered to be part of a sentence and should be punctuated accordingly. In the above example, a comma follows the equation because the next line is a subordinate clause.  If the equation ends the sentence, a period should follow the equation.  The line following an equation should not be indented unless it is meant to start a new paragraph.  Indentation after an equation is avoided in LaTeX by not leaving a blank line between the equation and the subsequent text.

References to equations include the equation number in parentheses, for example, ``Equation~(\ref{eq:alpha}) shows ..." or ``Combining Eqs.~(2) and (3), we obtain..."  Using a tilde in the LaTeX source file between two characters avoids unwanted line breaks.

%%-----------------------------------------------------------
\subsection{Formatting Theorems}

To include theorems in a formal way, the theorem identification should appear in a 10-point, bold font, left justified and followed by a period.  The text of the theorem continues on the same line in normal, 10-point font.  For example,

\noindent{\bf Theorem 1.} For any unbiased estimator...

Formal statements of lemmas and algorithms receive a similar treatment.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{SOME LATEX GUIDANCE} \label{sec:latex}

%%-----------------------------------------------------------
\subsection{Margins and PostScript Fonts}

Manuscripts submitted electronically to as PostScript (PS) files must have the correct margins. LaTeX margins are related to the document's paper size. The paper size is set at two separate places in the process of creating a PS file. The first place is in {\tt latex}. The default in {\tt article.tex}, on which {\tt spie.cls} is based, is USA letter paper. To format a document for A4 paper, the first line of the LaTeX source file should be \verb|\documentclass[a4paper]{spie}|.

The output of the LaTeX utility is a file with the extension DVI (for Device Independent), which encodes the formatted document.  The application DVIPS is typically used to convert the DVI file to a PS file.  DVIPS has its own default paper size, which can be overridden with the option ``{\tt -t letter}" or ``{\tt -t a4}".  
If the foregoing steps do not produce the correct top margin, you can move the text lower on the page (by 9 mm) with the command \verb|\addtolength{\voffset}{9mm}|, placed right after the \verb|\documentclass| command, for example.

Another DVIPS option specifies the incorporation of (scalable) PostScript Type 1 fonts in its output PS file. This feature is important for obtaining a subsequent PDF file that will be clearly displayed on a computer monitor by Adobe Acrobat Reader.  The option ``{\tt -P pdf}" makes DVIPS include these fonts in its output PS file.

%%-----------------------------------------------------------
\subsection{Bold Math Symbols}

The math package from the American Mathematical Society allows one to easily produce bold math symbols, well beyond what is available in LaTeX. It also provides many useful capabilities for creating elaborate mathematical expressions. You need to load the AMS math package near the top of the LaTeX source file, right after the \verb+\documentclass+ command:\\[1ex]
\verb+\usepackage[]{amsmath}+ \\[1ex]
Then for bold math symbols use \verb+\boldsymbol+ in equations, e.g., 
\verb+$\boldsymbol{\pi}$+ 
yields a bold pi.  You can make it easier to use by defining a command:\\[1ex]
\verb+\newcommand{\bm}[1]{\boldsymbol{#1}}+ \\[1ex]
and then using it like so \verb+$\bm{\pi}$+.

Not all math symbols are available in bold.  In a pinch, you can use \verb+\pmb+ ("poor man's bold"), which is defined in \verb+amsmath+. This command approximates a bold character with a superposition of several, slightly displaced unbold characters.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{abstract}
  Automated 3D modeling of building interiors is used in applications
  such as virtual reality and environment mapping. Texturing these
  models allows for photorealistic visualizations of the data
  collected by such modeling systems. Camera poses obtained by these
  systems often suffer from inaccuracies, resulting in visible
  discontinuities when successive images are projected adjacently onto
  a surface for texturing. Existing methods to stitch images together
  are often computationally expensive and work independently of pose
  estimates and geometry information. We propose a method to refine
  camera poses using both existing estimates and geometry information,
  followed by two different methods to composite images together,
  based on the uniformity of available images. The effectiveness of
  our methods is demonstrated on a number of different indoor
  environments.
\end{abstract}

@@

The first step in automated 3D modeling is the physical scanning of
an environment's geometry. An indoor modeling system must recover its
own poses within an environment while simultaneously reconstructing
the 3D structure of the environment itself. CITATION. This is known as
the simultaneous localization and mapping (SLAM) problem, and is
generally solved by taking readings from laser range scanners,
cameras, and inertial measurement units (IMUs) at multiple locations
within the environment.

In this paper, we work with data obtained from a backpack-mounted
system, carried by an ambulatory human. Such a system provides
advantages over more common wheel-mounted systems in terms of agility
and portability, but results in much higher localization error and
uncertainty overall. CITATION. As a result, common methods for texture
mapping generally produce poor results, leading to the development of
the approaches contained in this paper. Besides localization
information and images, environment geometry is required as well for
texture mapping. In this paper we work mainly with models obtained
by fitting planar surfaces to point clouds generated by the
backpack system. CITATION. We have also developed an appropriate
method to work with near-planar triangular meshes, as described in APPENDIX
SOMETHING.

Though our texture mapping procedure was designed with the
aforementioned system in mind, it is easily generalizable to any other
system, provided the required inputs of planar geometry, images with
rough extrinsic pose estimates, and either known or estimated
intrinsic camera calibration matrices. An important benefit of our method
is its relatively low complexity and modularity of each step,
allowing for quick tuning of parameters and easy accomodation of new
input data given updates in localization or model-generation
algorithms.

The remainder of the paper is organized as follows. Section X explains
how input images are projected onto our geometry and simple approaches
towards texturing. Section X covers more sophisticated existing
approaches to image matching and stitching, and their performance on
our datasets. Section X contains our approach towards image alignment,
followed by Section X, which describes compositing approaches. Section
X contains further results and conclusions.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

Once we have a list of candidate images for $t$, we define a scoring
function in order to objectively select the best image for texturing
$t$. Since camera pose errors compound over distance, we wish to
minimize the distance between cameras and the surfaces they
texture. Additionally, we desire images that are projected
perpendicularly onto the plane, maximizing the resolution and amount
of useful texture available in their projections, as well as
minimizing any parallax effects due to real-world geometry not
accurately represented by our digital model. In other words, we wish
to minimize the angle between the tile's normal vector and the camera
axis for images selected for texture mapping. These two criteria can
be met by maximizing the function $\frac{1}{d} (-1 \cdot \vec{c})
\cdot \vec{n}$ as shown in Figure
\ref{fig:scoringFunction}. Specifically, $d$ is the distance between
the centers of a camera and a tile, and $\vec{n}$ and $\vec{c}$ are
the directions of the plane's normal and the camera axis respectively.

@@

As Figure \ref{fig:compareAll}(a) demonstrates, this approach leads to
the best texture for each tile independently, but overall results in
many image boundaries with abrupt discontinuities, due to significant
misalignment between images. Given our high number of input images, one approach might be to more intelligently only select images that match up well, but the far more robust solution is to match up images directly such that we maximize our amount of useful images.

@@

\section{Efficient Image Alignment with Geometry Information}

In this section, we describe our method of efficient and robust image
alignment. Our approach consists of four parts. First, all images are
projected onto the surface and rotated such that detected lines within
them are parallel to lines comprising the surface's boundary and
intersection with other surfaces. Next, these projections are shifted
such that interior lines near surface intersections are matched
up. Following this, occlusion checks are performed across each image
to remove invalid parts of each image for the target surface. These
first three steps are all image-independent and can be considered
pre-processing steps and performed in parallel. Finally, we detect
SIFT feature matches between pairs of images and solve a linear least
squares problem in 2D to maximize matches.

\subsection{Geometry-based Alignment}
\label{sec:projectionAndRotation}
After computing each image's projection onto the target surface, as
described in Section \ref{sec:simpleTextureMapping}, we can detect
lines in the projections using Hough transforms. Experience and
intuition show that walls in indoor environments contain linear
features that are either horizontal or vertical, often corresponding
to doors, windows, posters, etc. Thus, when texturing walls, we rotate
images in 2D such that dominant lines are made to be horizontal or
vertical. Though not as relevant in our datasets, this can be extended
such that lines are oriented with a wall's boundaries, for example in
areas with a slanted roof.

At this point in time, image occlusions have not been accounted
for. As a result, some image projections contain texture that should
project to an adjacent surface, generally with a linear boundary where
the two surfaces meet. If this linear boundary is detected by Hough
transform, it makes sense to rotate and align this boundary such that
the visual boundary between two surfaces matches the physical boundary
in our digital model.

To perform the rotation and alignment, we do MAGIC.  Then we indicate
that magic has been done so that in the next step stuff is good. Mark
as anchor images.

\subsection{Image Occlusion}
get from github i guess i dunno. or check avz email

\subsection{2D Feature Alignment}
\label{sec:robustSIFTFeatureMatching}
Our next step is to align overlapping images by searching for
corresponding points between all pairs of overlapping images. We use
SIFT features for their high detection rate, and choose to use feature
alignment rather than intensity-based alignment due to the differences
in lighting as well as possible occlusion among our images, both of
which feature alignment is less sensitive to \cite{lai1999robust,
  lowe1999object, mikolajczyk2005performance, szeliski2006image}.
SIFTGPU MORE CITATION SIFT matches determine $d^x$ and $d^y$ distances
between each pair of features for two images on the plane, though
these distances may not always be the same for different
features. Since indoor environments often contain repetitive features
such as floor tiles or doors, we need to ensure that SIFT-based
distances are reliable. In order to mitigate the effect of incorrect
matches and outliers, the RANSAC framework \cite{fischler1981random}
is used for a robust estimate of the optimal $d^x_{i,j}$ and
$d^y_{i,j}$ distances between two images $i$ and $j$. The RANSAC
framework handles the consensus-building machinery, and requires a
fitting function and a distance function. For this application, the
fitting function simply finds the average distance between matches in
a pair of images. The distance function for a pair of points is chosen
to be the difference between those points' SIFT match distance and the
average distance computed by the fitting function. We use a 10 pixel
outlier threshold, so that SIFT matches are labeled as outliers if
their horizontal or vertical distances are not within 10 pixels of the
average distance computed by the fitting function.

We now use the $d^x_{i,j}$ and $d^y_{i,j}$ distances between each pair
of images to refine their positions using least squares. Recall that
there are a total of $M^{2}$ possible pairs of images, though we only
generate distances between images that overlap at SIFT feature
points. Given these distances and the original image location
estimates, we can solve a least squares problem
($\textrm{min}_{\vec{\beta}} ||A \vec{\beta} - \vec{\gamma}||_2^2 $)
to estimate the location of the images on the plane. The
$M$-dimensional vector $\vec{\beta}$ represents the unknown $x$
location of each image on the plane for $1 \dots M$. The optimal $x$
and $y$ locations are obtained in the same way, so we only consider
the $x$ locations here:

\[\vec{\beta} =
\begin{pmatrix}
  x_1, & x_2, & x_3, & \cdots & x_{M-1}, & x_M
\end{pmatrix}
\]

The $N \times (M+1)$ dimensional matrix $A$ is constructed with one
row for each pair of images with measured distances produced by the
SIFT matching stage. A row in the matrix has a $-1$ and $1$ in the
columns corresponding to the two images in the pair. For example, the
matrix below indicates a SIFT-based distance between images 1 and 2,
images 1 and 3, images 2 and 3, etc.
\[
A =
\begin{pmatrix}
  -1 & 1 & 0 & \cdots & 0 & 0\\
  -1 & 0 & 1 & \cdots & 0 & 0\\
  0 & -1 & 1 & \cdots & 0 & 0\\
  \vdots  & \vdots & \vdots & \ddots & \vdots  & \vdots\\
  0 & 0 & 0 & \cdots & 1 & 0 \\
  0 & 0 & 0 & \cdots & -1 & 1 \\
  1 & 0 & 0 & \cdots & 0 & 0 \\
\end{pmatrix}
\]
If only relative distances between images are included then there is
no way to determine the absolute location of any of the images and the
matrix becomes rank deficient. To fix this we choose the first image
to serve as the anchor for the rest, meaning all the absolute
distances are based on its original location. This is done by adding a
row with a $1$ in the first column and the rest zeros.

Finally, the $N$-dimensional observation vector $\vec{\gamma}$ is
constructed using the SIFT-based distances generated earlier in the
RANSAC matching stage. The last element in the observation vector is
the location of the first image determined by its original noisy
localization, from \cite{chen2010indoor, liu2010indoor}. Thus
$\vec{\gamma}$ can be written as:

\[
\vec{\gamma}^T =
\begin{pmatrix}
  d_{1,2}, &d_{1,3}, &d_{2,3}, &\hdots &d_{N-2,N-1}, &d_{N-1,N}, &x_1
\end{pmatrix}
\]

The $\vec{\beta}$ that minimizes $||A \vec{\beta} -
\vec{\gamma}||_2^2$ results in a set of image locations on the plane
that best honors all the SIFT-based distance measurements between
images. In practice there is often a break in the chain of images,
meaning that no SIFT matches are found between one segment of the
plane and another. In this case we add rows to the $A$ matrix and
observations to the $\vec{\gamma}$ vector that contain the original
noisy $x$ and $y$ distance estimates from the original camera
poses. Another way to do this is to add rows for all neighboring pairs
of images and solve a weighted least squares problem where the SIFT
distances are given a higher weight i.e. 1, and the noisy distances
generated by the original camera poses \cite{chen2010indoor,
  liu2010indoor} are given a smaller weight e.g. 0.01.

After completing this same process for the $y$ dimension as well, and
making the resultant shifts, our images overlap and match each other
with far greater accuracy. Applying the tile caching method from
Section \ref{sec:mappingWithCaching} on these rotated and shifted
images results in the significant improvements shown in Figure
\ref{fig:compareAll}(c), as compared to Figure
\ref{fig:compareAll}(b).

\section{Image Compositing}

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=1in]{projectionWall.pdf}}
  \centering
  \subfloat[][]{\includegraphics[width=1.1in]{projectionCeiling.pdf}}
  \centering
  \subfloat[][]{\includegraphics[width=0.9in]{projectionWallCrop.pdf}}
  \caption{(a) Undistorted fisheye images for vertical planes are
    tilted, but their effective camera axis is more or less normal to
    the plane. (b) The effective camera axis for ceilings is at an
    angle with respect to the plane normal. (c) Wall images are
    cropped to be rectangular.}
  \label{fig:projectionAngles}
\end{figure}

As mentioned earlier, the mobile backpack system used to capture data
in this paper contains 2 cameras with $180^\circ$ fisheye lenses
facing to the right and left. Since most human operators do not carry
the backpack in a perfectly upright position and are bent forwards at
15 to 20 degrees with respect to the vertical direction, the
undistorted fisheye images are head on with respect to vertical walls,
but at an angle with respect to horizontal floors and ceilings. This
is depicted for wall and ceiling planes in Figures
\ref{fig:projectionAngles}(a) and \ref{fig:projectionAngles}(b)
respectively. These oblique camera angles for ceilings translate into
textures that span extremely large areas once projected, as shown in
Figure \ref{fig:projectionAngles}(b). Using the tile-based texture
mapping criteria from Figure \ref{fig:scoringFunction}, such
projections have highly varying scores depending on the location on
the plane, as shown in Figure \ref{fig:projectionAngles}(b). Thus, the
tiling approach in Section \ref{sec:simpleTextureMapping} is a
reasonable choice for texturing floors and ceilings, as it uses the
parts of image projections that are viable choices for their
respective plane locations, e.g. areas near point A in Figure
\ref{fig:projectionAngles}(b), but not near point B.

\subsection{Tile-Mapping with Caching}
\label{sec:mappingWithCaching}
Since discontinuities occur where adjacent tiles select non-aligned
images, it makes sense to take into account image selections made by
neighboring tiles while texture mapping a given tile. By using the
same image across tile boundaries, we can eliminate a discontinuity
altogether. If this is not possible because a tile is not visible in
images used by neighboring tiles, using similar images across tile
boundaries also leads to less noticeable discontinuities.

\subsection{Shortest}
For wall planes however, most images are taken from close distances
and more or less head-on angles, resulting in higher resolution
fronto-parallel projections. As a result, for each tile on a wall
plane, the scoring function of Figure \ref{fig:scoringFunction} is
relatively flat with respect to a large number of captured images, as
they are all more or less head on. Thus, the scoring function is less
significant for walls, and it is conceivable to use a different
texturing strategy for them so as to minimize visible seams in texture
mapped planes. This can be done by choosing the smallest possible set
of images that (a) covers the entire plane and (b) minimizes the
amount of visible seams between them. A straightforward cost function
that accomplishes the latter is the Sum of Squared Differences (SSD)
of pixels in overlapping regions between all pairs of images, after
they have been rotated as described in Section
\ref{sec:projectionAndRotation}. Minimizing this cost function
encourages image boundaries to occur either in featureless areas, such
as bare walls, or in areas where images match extremely well.

To cover the entirety of a plane, our problem can be defined as
minimally covering a polygon i.e. the plane, using other polygons of
arbitrary geometry i.e. image projections, with the added constraint
of minimizing the cost function between chosen images.  This is a
complex problem, though we can take a number of steps to simplify it.

Given that wall-texture candidate images are taken from more or less
head-on angles, and assuming only minor rotations are made in Section
\ref{sec:projectionAndRotation}, we can reason that their projections
onto the plane are approximately rectangular. By cropping them all to
be rectangular, as shown in Figure \ref{fig:projectionAngles}(c), our
problem becomes the conceptually simpler one of filling a polygon with
rectangles, such that the sum of all costs between each pair of
rectangles is minimal. We thus also retain the advantages of working
with rectangular units, as explained in Section
\ref{sec:simpleTextureMapping}.

The location and orientation of the cameras and the fisheye lenses on
the acquisition backpack is such that images nearly always contain the
entirety of the floor to ceiling range of wall planes. Images are
therefore rarely projected with one above another when texturing wall
planes. In essence, we need only to ensure lateral coverage of wall
planes, e.g. from left to right, as our images provide full vertical
coverage themselves. We can thus construct a Directed Acyclic Graph
(DAG) from the images, with edge costs defined by the SSD cost
function, and solve a simple shortest path problem to find an optimal
subset of images with regard to the cost function \cite{dijkstra}.

\begin{figure}
  \centering
  \includegraphics[width=3in]{dagCreation.pdf}
  \caption{DAG construction for the image selection process. \\}
  \label{fig:dagCreation}
\end{figure}

Figure \ref{fig:dagCreation} demonstrates the construction of a DAG
from overlapping images of a long hallway. Images are sorted by
horizontal location left to right, and become nodes in a
graph. Directed edges are placed in the graph from left to right
between overlapping images. The weights of these edges are determined
by the SSD cost function. Next, we add two artificial nodes, one start
node representing the left border of the plane, and one end node
representing the right border of the plane. The left(right) artificial
node has directed edges with equal cost $C_0$ to(from) all images that
meet the left(right) border of the plane.

We now solve the shortest path problem from the start node to the end
node. This results in a set of images completely covering the plane
horizontally, while minimizing the cost of seams between images.

In rare cases where the vertical dimension of the plane is not
entirely covered by one or more chosen images, we are left with holes
where no images are selected to texture. To address this, we solve for
a new shortest path after modifying the edges chosen by the original
shortest path solution such that they have higher cost than all other
edges in the DAG, and solve for a new shortest path. This ensures that
these edges are not chosen again unless they are the only available
option. Our second shortest path solution results in a new set of
chosen images, of which we only retain those that cover areas not
covered by the first set. This process can be repeated as many times
as needed, until holes are no longer present. This method is not as
optimal as a 2D-coverage solution would be, but it is a fast
approximation, and adequately handles the few holes we encounter.

With this completed, we have now mapped every location on the plane to
at least one image, and have minimized the number of images, as well
as the discontinuities between their borders. As seen in Figure
\ref{fig:compareAll}(d), this seam minimization method has fewer
visible discontinuities than Figure \ref{fig:compareAll}(c)
corresponding to the tile caching approach\footnote{In Figure
  \ref{fig:compareAll}(d), we arbitrarily chose one image for
  texturing where images overlap, as blending will be discussed in the
  next section.}. This is especially evident when comparing the
posters in the images, which have clear misalignment over seams in
Figure \ref{fig:compareAll}(c), but are much more aligned in Figure
\ref{fig:compareAll}(d). This is because the seam minimization
approach directly reduces the cost of each image boundary, while the
tile caching method uses a scoring function that only approximates
this effect. Furthermore, seam minimization guarantees the best
selection of images, while the sequential tile caching method may
select images early on that turn out to be poor choices once
subsequent tiles have been processed. The seam minimization approach
is also far less intensive in terms of memory usage and runtime, both
during texture generation and rendering, as it does not require
discretizing planes or images\footnote {For the seam minimization
  approach, occlusion checks are performed over the entirety of each
  image to determine available areas for texture mapping. Since indoor
  environments mostly consist of vertical or horizontal surfaces with
  high amounts of right angles, image projections remain largely
  rectangular after masking out occluded areas. Therefore, as a
  preprocessing step, we can recursively split each image into
  rectangular pieces and check for occlusions in a similar manner as
  in the tiling process.  }.
